Report

### 2022-05-12 manuscript.Rmd

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [REF].
Replication research has come into focus as a solution to this crisis
and a way to derive knowledge that will stand the test of time.
In step with this interest in replication research,
research synthesis methods have blossomed.
These methods aggregate the findings of multiple studies,
and thus enable drawing overarching conclusions across multiple studies.
Both quantitative [REF] and qualitative [REF TMSR] research synthesis methods exist, but the present paper focuses on quantitative methods.

A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity.
Heterogeneity appears when studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
Meta-analysis is the most common research synthesis method.
In meta-analysis, heterogeneity can be accounted for in four ways.
First, if studies are exact replications, one can assume that no heterogeneity exists and conduct a fixed-effect meta-analysis to estimate their common population effect.
Second, when differences between studies can be assumed to be random, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects.
Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
And finally, when there are many potential differences and it is not known beforehand which are relevant,
exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
Each of these approaches requires making different assumptions about the nature of heterogeneity.

An alternative approach that does not impose such assumptions is Bayesian evidence synthesis [BES, REF Kuiper 2013].
BES focuses on the aggregation of evidence in favor of an informative hypothesis $H_i$ across studies.
The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
A Bayes factor can be interpreed as the ratio of evidence in favor of $H_i$ divided by evidence against it.
Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it.
These Bayes factors can be synthesized across studies by taking their product.
The resulting product Bayes factor PBF summarizes the total evidence for the hypothesis.
Note that other approaches to BES exist; for an overview, see REF Klugkist.

This tutorial paper introduces the first implementation of BES in userfriendly open source software.
A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version XXX.
This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
and illustrates several use cases through reproducible examples.

@@

To examine performance in a range of realistic scenarios,
several design factors were manipulated:
The population effect size $\rho \in (-0.1, 0, 0.1, 0.2)$,
the sample size per group $n \in (20, 80, 200, 500)$,
the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents an assumption of perfect reliability, as is the case when analyzing correlations between observed items or mean scale scores.
The design factors combined to produce `r out$conditions` unique conditions.
For all simulation conditions, 1000 data sets were generated.

@@

knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
```
PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and residual heterogeneity.

@@

Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increase with sample size.
The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
This difference explains the effect of errorsd on the diference between algorithms (see Table \@ref(tab:tabeffect)).

@@

### Effect of residual heterogeneity

```{r, fig.cap="Mean performance by errorsd"}

@@

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- df_plot[ ,list(mean=mean(val)), by=c("errorsd", "Outcome", "alg")]
ggplot(df_plot, aes(x = errorsd, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance")
```

The figure indicates that for PBF, at higher levels of residual heterogeneity, lower sensitivity is exchanged for greater specificity.
The other algorithms do not show this pattern, as their specificity is at a ceiling.
Their sensitivity decreases with residual heterogeneity,
and therefore, so does their overall performance.
This difference in pattern of effects explains why residual heterogeneity has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Note that, in contrast to other algorithms, the performance of PBF is less susceptible to residual heterogeneity.

@@

The overall performance of PBF increased most with increasing sample size.
With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.
With increasing residual heterogeneity, increasing specificity was traded off for decreasing sensitivity.

>>>>

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [REF].
Replication research has come into focus as a solution to this crisis
and a way to derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021].
In step with this interest in replication research,
research synthesis methods have blossomed [REF].
These methods aggregate the findings of multiple studies,
and thus enable drawing overarching conclusions across multiple studies.
Both quantitative [e.g., @vanlissaSelectRelevantModerators2021] and qualitative [e.g., @vanlissaMappingPhenomenaRelevant2021] research synthesis methods exist.
The present paper focuses on quantitative methods.

A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
Heterogeneity appears when, for example, studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
Meta-analysis is the most common research synthesis method [@borensteinIntroductionMetaAnalysis2009].
In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
First, if studies are exact replications, one can assume that no heterogeneity exists and conduct a fixed-effect meta-analysis to estimate their common population effect.
Second, when differences between studies can be assumed to be random, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects.
Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
And finally, when there are many potential differences and it is not known beforehand which are relevant,
exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
Each of these approaches requires making different assumptions about the nature of heterogeneity.

An alternative approach that does not impose such assumptions is Bayesian evidence synthesis [BES, @kuiperCombiningStatisticalEvidence2013].
BES focuses on the aggregation of evidence in favor of an informative hypothesis $H_i$ across studies.
The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
A Bayes factor can be interpreed as the ratio of evidence in favor of $H_i$ divided by evidence against it.
Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it.
These Bayes factors can be synthesized across studies by taking their product.
The resulting product Bayes factor PBF summarizes the total evidence for the hypothesis.
Note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].

This tutorial paper introduces the first implementation of BES in userfriendly open source software.
A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `r as.character(packageVersion("bain"))`.
This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
and illustrates several use cases through reproducible examples.

@@

To examine performance in a range of realistic scenarios,
several design factors were manipulated:
The population effect size $\rho \in (-0.1, 0, 0.1, 0.2)$,
the sample size per group $n \in (20, 80, 200, 500)$,
the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is the assumed when analyzing correlations between observed items or scale scores.
The design factors combined to produce `r out$conditions` unique conditions.
For all simulation conditions, 1000 data sets were generated.

@@

knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
```
PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.

@@

Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increase with sample size.
The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
This difference explains the effect of reliability on the diference between algorithms (see Table \@ref(tab:tabeffect)).

@@

### Effect of reliability

```{r, fig.cap="Mean performance by reliability"}

@@

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- df_plot[ ,list(mean=mean(val)), by=c("errorsd", "Outcome", "alg")]
ggplot(df_plot, aes(x = errorsd, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Reliability")
```

The figure indicates that for PBF, at higher reliability, lower sensitivity is exchanged for greater specificity.
The other algorithms do not show this pattern, as their specificity is at a ceiling.
Their sensitivity decreases with lower reliability,
and therefore, so does their overall performance.
This difference in pattern of effects explains why reliability has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Note that, in contrast to other algorithms, the performance of PBF is less susceptible to reliability.

@@

The overall performance of PBF increased most with increasing sample size.
With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.
With increasing reliability, increasing specificity was traded off for decreasing sensitivity.
